TODO: A Survey of StarCraft AI Techniques

### 2013-02-15 survey.tex
Details about AIUR, and url for each bot.

As well as spatial reasoning, temporal reasoning is related to tactics
since it implies  to decide about army movements,  like when attacking
or withdrawing  units from battle.  Strategy also copes  with temporal
reasoning, mainly  with planning problems:  when researching upgrades,
when constructing wished buildings,  when switching strategy if needed,
etc.

@@

\begin{itemize}
\item BroodwarBotQ: uses abstraction for micro-management, and divide-and-conquer for macro-management and intelligence gathering. To avoid conflicts between modules (since the individual tasks of each of the modules are not completely independent), BBQ uses an arbitrator.
\item Nova: is similar in design as BroodwarBotQ, and uses abstraction for micro-management, and divide-and-conquer for macro-management. The differences are that Nova does not have an arbitrator to resolve conflicts, but has a higher-level module ({\em strategy manager}), which posts information to the blackboard that the rest of modules follow (thus, making use of abstraction).
\item UAlbertaBot: also uses abstraction in micro-management like the previous two bots. But it also uses it in macro-management: as can be seen, the production manager sends commands to the building manager, who is in charge of producing the buildings. This bot also uses divide-and-conquer, and tasks like scouting and resource gathering are managed by separate, independent modules.
\item Skynet: makes extensive use of both abstraction and divide-and-conquer. Considering the decision making component of Skynet, we can see a high level module that issues commands to a series of tactics modules. The collection of tactic modules queue {\em tasks} (that are analogous to the abstract actions used in SPAR). Each different task has a specific low level module that knows how to execute it. Thus, Skynet uses a 3 layered abstraction hierarchy, and uses divide-and-conquer in all levels except the highest.
\item SPAR: only uses abstraction. Its high-level module determines the strategy to use, and the tactical decision module divides it into a collection of {\em abstract actions}, that are executed by the lower-level modules.
\item AIUR:  is mainly divide-and-conquer oriented, for  macro as well
  as  for  micro,  with  a  slight  abstraction  on  macro  due  to  a
  SpendManager deciding how to spend and share resources among Base, Production and Construction Managers. At the beginning of a game, the MoodManager initializes a ``mood'' which will influence both micro and macro. For example, micro-management is divided into three independent managers: the {\em Defense Manager}, controlling military units when there is nothing special, the {\em Under Attack Manager}, activated when the opponent is attacking our bases, and the {\em Army Manager}, taking control of units when it is time to attack, following a timing given by the current mood.
\item BTHAI: uses a two-tier abstraction hierarchy, where a collection of high-level modules command a collection of lower-level agents in charge of each of the units. At the high-level, BTHAI uses divide-and-conquer, having multiple high-level modules issuing commands to the lower-level units.
\end{itemize}

@@

Some bots using divide-and-conquer, assume that each of the modules can act independently and that their actions can be executed without interference. BBQ, UAlbertaBot and AIUR, however use an arbitrator ({\em Game Commander'} in UAlbertaBot) that makes sure that modules do not send contradictory orders to the same unit. However, very little bots handle the problem of how to coordinate resource usage amongst modules, for instance BTHAI uses a first-come-first-serve policy for spending resources, the first module that requests resources is the one that gets them. Nova and Skynet are exceptions, and implement some rudimentary prioritization based on the high level strategy. Following available resources and timing, AIUR's {\em Spend Manager} orders Base, Production and Construction Managers what they have to build/produce. It also orders to start tech research and upgrades.

One interesting aspect of the seven bots described above is that, while all of them are reactive at the lower level (reactive control), most if not all of them, are scripted at the highest level of abstraction. BTHAI reads build and squad formations from a predefined script, Nova's {\em Strategy Manager} is a predefined finite-state machine, BBQ's construction manager reads the build order from a predefined script, and Skynet's {\em BuildOrder Manager} is basically a predefined script. Such scripts describe the strategy that the bots will use, however, such strategy is always fixed. One could see this pre-scripting as if each bot defined a ``high-level programming language'' to describe Starcraft strategies, and the bots themselves are just interpreters of such strategy. Compared to current approaches for Chess or Go, this scripting seems a rigid and inflexible, but responds to the much higher complexity of the Starcraft game. An interesting exception to that is UAlbertaBot, which uses a search algorithm in the {\em Production Manager} to find near-optimal build orders. Another interesting case is AIUR, that uses a {\em Mood Manager} to randomly pick a mood among five (rush, aggressive, defensive, macro, fast expand), which will influence the build order, strategy and tactics, all of them scripted so far. This mood can change during the game, following the opponent behavior.

@@

\begin{table}[bth]
\caption{Results of the AIIDE 2010 competition.}
\label{tab:aiide2010}
\begin{small}
\begin{center}
% bracket A
\begin{tabular}{|l|c|}
\hline
Position & bot\\ \hline
1 & Overmind \\
2 & Krasi0 \\
3 & Chronos \\ \hline
\end{tabular}
\end{center}
\end{small}
\end{table}

@@

In the end, Skynet\footnote{\url{http://code.google.com/p/skynetbot/}} won the competition with its solid Protoss play (results are summarized in Table \ref{tab:aiide2011}). The bot executed one of a small set of
strategies randomly at the start of the match based on the map and the race of the opponent. Skynet would
then amass a medium to large sized army and expand before moving out to attack. Good use of Dragoon range
and kiting micromanagement allowed it to hold off the early aggression of other bots such as UAlbertaBot,
which came in 2nd.

@@

AIUR also chose Protoss, with a strategy which was in between Skynet and UAlbertaBot in terms of attack
timings. Of note in these results was that a rock-paper-scissors effect happened among the top 3 finishers.
Of the 30 rounds, Skynet beat UAlbertaBot 26 times, UAlbertaBot beat AIUR 29 times, and AIUR beat Skynet
19 times. Another notable result is that Overmind did not choose to compete despite winning the 2010
competition. After the competition, many bot programmers (including the Overmind team) realized that 
their 2010 strategy was quite easily defeated by early game rushing strategies, and so they submitted
a Terran bot instead, called Undermind, which finished in 7th.

\begin{table}[bth]
\caption{Results of the AIIDE 2011 competition.}
\label{tab:aiide2011}
\begin{small}
\begin{center}
% bracket A
\begin{tabular}{|l|c|c|}
\hline
Position & bot & win \% \\ \hline
1 & Skynet & 88.9\% \\
2 & UAlbertaBot & 79.4\% \\
3 & AIUR & 70.3\% \\
4 & ItayUndermind & 65.8\% \\
5 & EISBot & 60.6\% \\ \hline
\end{tabular}
\end{center}
\end{small}
\end{table}

@@

Skynet once again won the competition with its solid Protoss build orders and good Dragoon kiting. AIUR
and UAlbertaBot switched positions from the previous year to come 2nd and 3rd respectively. Both AIUR
and UAlbertaBot used data stored from the results of previous games to select a strategy for future matches.
UAlbertaBot did this using the UCB algorithm, while AIUR used a custom algorithm to achieve similar
results. The effect of this strategy selection process can be seen Figure \ref{fig:aiide2012} which shows bot
win percentages over time. While the earlier rounds of the tournament fluctuated wildly in results,
eventually the results converged to their final values. One of the main reasons for this is due to
the bots learning which strategies to use as the tournament progressed.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/aiide2012.png}
    \caption{Evolution of the win percentage of each bot participating in the AIIDE 2012 competition}
    \label{fig:aiide2012}
\end{figure}

\begin{table}[bth]
\caption{Results of the AIIDE 2012 competition.}
\label{tab:aiide2012}
\begin{small}
\begin{center}
% bracket A
\begin{tabular}{|l|c|c|}
\hline
Position & bot & win \% \\ \hline
1 & Skynet & 84.4\% \\
2 & AIUR & 72.2\% \\
3 & UAlbertaBot & 68.6\% \\
4 & BroodwarBotQ & 59.1\% \\
5 & AdjutantBot & 52.8\% \\ \hline
\end{tabular}
\end{center}
\end{small}
\end{table}

@@

\begin{table*}[htb]
\caption{Results of the first round at CIG 2011, held in two brackets.
Qualified for the final round: UAlbertaBot and Skynet (from A), Xelnaga
and BroodwarBotQ (from B, the latter by comparing direct encounters
with BTHAI of which 6:4 were won).}
\label{tab:cig-first-round}
\begin{small}
\begin{center}
% bracket A
\begin{tabular}{|l|c|c|c|}
\hline
\multicolumn{4}{|c|}{Bracket A} \\ \hline
Crashes & games & bot	& wins\\ \hline
0 &	 40 &	 UAlbertaBot &	 33\\
1 &   40 &	 Skynet	  &  31\\
2 &	 40 &	 AIUR	  &  24\\
1 &	 40 &	 Nova	  &  8\\
0 &	 40 &	 LSAI	  &  4\\
\hline
\end{tabular}
% bracket B
\begin{tabular}{|l|c|c|c|}
\hline
\multicolumn{4}{|c|}{Bracket B} \\ \hline
Crashes & games & bot	& wins\\ \hline
12 &	 40 &	 Xelnaga &	 25\\
3 &   40 &	 BroodwarBotQ  &  23 (6)\\
0 &	 40 &	 BTHAI	  &  23 (4)\\
17 &	 40 &	 Protoss Beast Jelly  & 17\\
0 &	 40 &	 EvoBot	  &  12\\
\hline
\end{tabular}
\end{center}
\end{small}
\end{table*}

@@

\begin{table}[bth]
\caption{Results of the CIG 2011 final round.}
\label{tab:cig-final-round}
\begin{small}
\begin{center}
% bracket A
\begin{tabular}{|l|c|c|c|}
\hline
Crashes & games & bot	& wins\\ \hline
0 &  30 &	 Skynet	  		&  26\\
0 &	 30 &	 UAlbertaBot 	&  22\\
3 &	 30 &	 Xelnaga  		&  11\\
2 &	 30 &	 BroodwarBotQ  	&  1\\
\hline
\end{tabular}
\end{center}
\end{small}
\end{table}

>>>>

As well as spatial reasoning, temporal reasoning is related to tactics
since it implies  to decide about army movements,  like when attacking
or  withdrawing  units from  battle.   A  part  of temporal  reasoning
consists  of  threat predictions,  where  one  has  to plan  when  the
opponent is  likely running an  attack.  Strategy copes  with temporal
reasoning  as well,  mainly with  planning problems:  when researching
upgrades, when constructing  wished buildings, when switching strategy
if needed, etc.

@@

\begin{itemize}
\item
  BroodwarBotQ\footnote{\url{http://github.com/SnippyHolloW/BroodwarBotQ}}:
  uses  abstraction for  micro-management, and  divide-and-conquer for
  macro-management  and  intelligence  gathering. To  avoid  conflicts
  between modules (since  the individual tasks of each  of the modules
  are not completely independent), BBQ uses an arbitrator.
\item  Nova\footnote{\url{http://nova.wolfwork.com/}}:  is similar  in
  design as  BroodwarBotQ, and uses  abstraction for micro-management,
  and  divide-and-conquer for  macro-management.  The differences  are
  that Nova does not have  an arbitrator to resolve conflicts, but has
  a  higher-level   module  ({\em  strategy   manager}),  which  posts
  information to the blackboard that the rest of modules follow (thus,
  making use of abstraction).
\item
  UAlbertaBot\footnote{\url{http://code.google.com/p/ualbertabot/}}:
  also  uses abstraction  in  micro-management like  the previous  two
  bots. But it  also uses it in macro-management: as  can be seen, the
  production manager sends commands to the building manager, who is in
  charge   of   producing   the   buildings.  This   bot   also   uses
  divide-and-conquer, and  tasks like scouting  and resource gathering
  are managed by separate, independent modules.
\item       Skynet\footnote{\url{http://code.google.com/p/skynetbot/}}:
  makes     extensive     use      of     both     abstraction     and
  divide-and-conquer.  Considering the  decision  making component  of
  Skynet, we  can see a  high level module  that issues commands  to a
  series of  tactics modules. The  collection of tactic  modules queue
  {\em  tasks} (that  are analogous  to the  abstract actions  used in
  SPAR).  Each different  task has  a specific  low level  module that
  knows how to  execute it. Thus, Skynet uses  a 3 layered abstraction
  hierarchy,  and uses  divide-and-conquer  in all  levels except  the
  highest.
\item
  SPAR\footnote{\url{http://www.planiart.usherbrooke.ca/projects/spar/}}:
  only uses abstraction. Its high-level module determines the strategy
  to  use,  and  the  tactical  decision  module  divides  it  into  a
  collection  of {\em  abstract  actions}, that  are  executed by  the
  lower-level modules.
\item  AIUR\footnote{\url{http://code.google.com/p/aiurproject/}}:  is
  mainly divide-and-conquer oriented, for  macro as well as for micro,
  with a  slight abstraction on  macro due to a  SpendManager deciding
  how  to  spend  and  share  resources  among  Base,  Production  and
  Construction Managers.  At the beginning of a  game, the MoodManager
  initializes  a  ``mood''  which  will  influence  both  tactics  and
  strategy. Tactics  are divided into three  independent managers: the
  {\em  Defense Manager},  controlling  military units  when there  is
  nothing special, the {\em  Under Attack Manager}, activated when the
  opponent is attacking our bases,  and the {\em Army Manager}, taking
  control of units when it is time to attack, following a timing given
  by the current mood.
\item  BTHAI\footnote{\url{http://code.google.com/p/bthai/}}:  uses  a
  two-tier  abstraction hierarchy,  where a  collection  of high-level
  modules command a collection of lower-level agents in charge of each
  of  the units.  At  the high-level,  BTHAI uses  divide-and-conquer,
  having   multiple  high-level  modules   issuing  commands   to  the
  lower-level units.
\end{itemize}

@@

Some bots  using divide-and-conquer, assume  that each of  the modules
can act independently  and that their actions can  be executed without
interference.  BBQ,  UAlbertaBot and  AIUR, however use  an arbitrator
({\em Game Commander'} in UAlbertaBot) that makes sure that modules do
not send contradictory orders to  the same unit.  However, very little
bots handle  the problem of  how to coordinate resource  usage amongst
modules, for  instance BTHAI uses a  first-come-first-serve policy for
spending resources,  the first module  that requests resources  is the
one that gets them. Nova and Skynet are exceptions, and implement some
rudimentary prioritization based on the high level strategy. Following
available  resources and  timing,  AIUR's {\em  Spend Manager}  orders
Base,  Production   and  Construction  Managers  what   they  have  to
build/produce. It also orders to start tech research and upgrades. The
idea here is  not to allocation resources among  different managers to
satisfy a need (informally: ``I want  to do that, so I need M minerals
and G gas''), but to do the  opposite, that is, to find how the AI can
spend the available money: ``I have right now M minerals and G gas, so
how can I spend it?''.

One  interesting aspect  of the  seven bots  described above  is that,
while all of them are  reactive at the lower level (reactive control),
most  if  not all  of  them,  are scripted  at  the  highest level  of
abstraction. BTHAI reads build  and squad formations from a predefined
script,  Nova's {\em  Strategy Manager}  is a  predefined finite-state
machine,  BBQ's construction  manager  reads the  build  order from  a
predefined script, and Skynet's  {\em BuildOrder Manager} is basically
a predefined script. Such scripts  describe the strategy that the bots
will use, however,  such strategy is always fixed.  One could see this
pre-scripting  as  if  each  bot defined  a  ``high-level  programming
language'' to  describe Starcraft strategies, and  the bots themselves
are just interpreters of such strategy. Compared to current approaches
for  Chess or Go,  this scripting  seems a  rigid and  inflexible, but
responds  to the  much higher  complexity  of the  Starcraft game.  An
interesting  exception to  that is  UAlbertaBot, which  uses  a search
algorithm in  the {\em Production Manager} to  find near-optimal build
orders.  Another interesting  case  is  AIUR, that  uses  a {\em  Mood
  Manager}  to   randomly  pick  a  mood  among   six  (cheese,  rush,
aggressive, defensive,  macro, fast expand), which  will influence the
build order,  strategy and tactics. All  of them scripted  so far, but
build  orders  can been  modified  but  the  Spend Manager,  following
available resources.  This mood can change during  the game, regarding
the opponent behavior.

@@

\begin{table}[t!]
\caption{Results of the AIIDE 2010 competition.}
\label{tab:aiide2010}
\begin{small}
\begin{center}
% bracket A
\begin{tabular}{|l|c|}
\hline
Position & bot\\ \hline
1 & Overmind \\
2 & Krasi0 \\
3 & Chronos \\ \hline
\end{tabular}
\end{center}
\end{small}
\end{table}

@@

In the  end, Skynet  won the competition  with its solid  Protoss play
(results  are  summarized   in  Table  \ref{tab:aiide2011}).  The  bot
executed one of a small set of strategies randomly at the start of the
match based on the map and the race of the opponent. Skynet would then
amass a  medium to large  sized army and  expand before moving  out to
attack. Good  use of Dragoon range and  kiting micromanagement allowed
it to hold off the early aggression of other bots such as UAlbertaBot,
which came in 2nd.

@@

AIUR also chose  Protoss, with a strategy which  was in between Skynet
and UAlbertaBot in terms of  attack timings.  At that time, AIUR chose
one  mood among  five (leading  to slightly  different  strategies and
tactics) at the  beginning of a game and kept it  until the end. These
five moods were:
\begin{itemize}
\item  Rush:  where  the  bot  tries  early  attacks,  and  have  good
  probabilities to send  the two or three first  Zealots to harass the
  opponent.
\item Aggressive: where  we have less chance to  perform harasses with
  the first  Zealots, and  the first attack  is usually a  bit delayed
  with regard to the Rush mood.
\item Macro: where the AI do not try any early attacks and focus a bit
  more on its economy before attacking.
\item Defense: where the AI  ``turtles'' and wait to have a consequent
  army before running an attack.
\item  Fast expand:  where the  first building  constructed it  a base
  expansion, for a very economical-oriented game.
\end{itemize}
Notice that  build orders are not  fully hard-coded since  they can be
altered by AIUR's Spend Manager.

Of  note  in  these  results  was that  a  rock-paper-scissors  effect
happened among  the top  3 finishers.  Of  the 30 rounds,  Skynet beat
UAlbertaBot 26  times, UAlbertaBot beat  AIUR 29 times, and  AIUR beat
Skynet  19 times.   Another notable  result is  that Overmind  did not
choose  to compete despite  winning the  2010 competition.   After the
competition,  many  bot  programmers  (including  the  Overmind  team)
realized that their  2010 strategy was quite easily  defeated by early
game rushing strategies,  and so they submitted a  Terran bot instead,
called Undermind, which finished in 7th.

\begin{table}[t!]
\caption{Results of the AIIDE 2011 competition.}
\label{tab:aiide2011}
\begin{small}
\begin{center}
% bracket A
\begin{tabular}{|l|c|c|}
\hline
Position & bot & win \% \\ \hline
1 & Skynet & 88.9\% \\
2 & UAlbertaBot & 79.4\% \\
3 & AIUR & 70.3\% \\
4 & ItayUndermind & 65.8\% \\
5 & EISBot & 60.6\% \\ \hline
\end{tabular}
\end{center}
\end{small}
\end{table}

@@

Skynet once  again won  the competition with  its solid  Protoss build
orders  and  good  Dragoon  kiting.   AIUR  and  UAlbertaBot  switched
positions from  the previous  year to come  2nd and  3rd respectively.
Both  AIUR  and UAlbertaBot  used  data  stored  from the  results  of
previous games  to select a strategy for  future matches.  UAlbertaBot
did  this  using  the  UCB   algorithm,  while  AIUR  used  a  uniform
distribution  to choose  its  mood before  altering this  distribution
after  some  games  against  the  same  opponent  to  favor  efficient
strategies, achieving  similar results than  UAlbertaBot. Notice that,
compared  to AIIDE 2011,  AIUR proposes  a new  mood, \textit{Cheese},
implementing a  Photon Cannon rush  strategy in order to  surprise the
opponent and  to finish the game  as soon as possible.   The effect of
this strategy selection process can be seen Figure \ref{fig:aiide2012}
which shows bot  win percentages over time.  While  the earlier rounds
of the tournament fluctuated wildly in results, eventually the results
converged to their  final values. One of the main  reasons for this is
due to  the bots  learning which strategies  to use as  the tournament
progressed.

\begin{figure}[t!]
    \centering
    \includegraphics[width=\columnwidth]{figures/aiide2012_v2}
    \caption{Evolution of the win percentage of each bot participating in the AIIDE 2012 competition}
    \label{fig:aiide2012}
\end{figure}

\begin{table}[t!]
\caption{Results of the AIIDE 2012 competition.}
\label{tab:aiide2012}
\begin{small}
\begin{center}
% bracket A
\begin{tabular}{|l|c|c|}
\hline
Position & bot & win \% \\ \hline
1 & Skynet & 84.4\% \\
2 & AIUR & 72.2\% \\
3 & UAlbertaBot & 68.6\% \\
4 & BroodwarBotQ & 59.1\% \\
5 & AdjutantBot & 52.8\% \\ \hline
\end{tabular}
\end{center}
\end{small}
\end{table}

@@

\begin{table*}[t!]
\caption{Results of the first round at CIG 2011, held in two brackets.
Qualified for the final round: UAlbertaBot and Skynet (from A), Xelnaga
and BroodwarBotQ (from B, the latter by comparing direct encounters
with BTHAI of which 6:4 were won).}
\label{tab:cig-first-round}
\begin{small}
\begin{center}
% bracket A
\begin{tabular}{|l|c|c|c|}
\hline
\multicolumn{4}{|c|}{Bracket A} \\ \hline
Crashes & games & bot	& wins\\ \hline
0 &	 40 &	 UAlbertaBot &	 33\\
1 &   40 &	 Skynet	  &  31\\
2 &	 40 &	 AIUR	  &  24\\
1 &	 40 &	 Nova	  &  8\\
0 &	 40 &	 LSAI	  &  4\\
\hline
\end{tabular}
% bracket B
\begin{tabular}{|l|c|c|c|}
\hline
\multicolumn{4}{|c|}{Bracket B} \\ \hline
Crashes & games & bot	& wins\\ \hline
12 &	 40 &	 Xelnaga &	 25\\
3 &   40 &	 BroodwarBotQ  &  23 (6)\\
0 &	 40 &	 BTHAI	  &  23 (4)\\
17 &	 40 &	 Protoss Beast Jelly  & 17\\
0 &	 40 &	 EvoBot	  &  12\\
\hline
\end{tabular}
\end{center}
\end{small}
\end{table*}

@@

\begin{table}[t!]
\caption{Results of the CIG 2011 final round.}
\label{tab:cig-final-round}
\begin{small}
\begin{center}
% bracket A
\begin{tabular}{|l|c|c|c|}
\hline
Crashes & games & bot	& wins\\ \hline
0 &  30 &	 Skynet	  		&  26\\
0 &	 30 &	 UAlbertaBot 	&  22\\
3 &	 30 &	 Xelnaga  		&  11\\
2 &	 30 &	 BroodwarBotQ  	&  1\\
\hline
\end{tabular}
\end{center}
\end{small}
\end{table}
